---
title: "Mini-Project 2: Do Incumbents Receive More Money?"
author: "Grace Hartley, Amrita Acharya, Callie Slevin"
date: "2019-03-24"
output: 
  html_document:
    code_folding: hide
---

^[https://github.com/hartlegr/sds192-mp2]

```{r, message=FALSE}
library(tidyverse)
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
devtools::install_github("benmarwick/wordcountaddin")

wordcountaddin::word_count("mp2.rmd")

```

We were interested in the sum of contributions made for and against the campaigns of different statuses of canidates (challenger, incumbent, and open seat canidates).

First, we joined the candidates table and the contributions table using the "cand_id."
```{r}
cand_contrib <- candidates%>% 
  left_join(contributions, "cand_id")
```


```{r, eval=FALSE}
library(Stat2Data)
tally(~transaction_type, data = cand_contrib)
```

Then, we filtered for the year 2012, while removing the rows with NA data. This provided us with all the contributions for canidates who last ran for office in 2012. We grouped the contributions by transaction type, and then mutated to add a column called 'for_cand'. This column provided information on whether the contribution was in favor of or in opposition of the canidate listed, listing either 'forcan' or 'agcan' respectively. 
```{r}
cand_contrib <- cand_contrib%>%
  filter(cand_election_yr == 2012)%>% 
  na.omit() %>%
  group_by(transaction_type)%>%
  mutate(for_cand = ifelse(transaction_type == "24K"|transaction_type ==  "24E", "forcan","agcan"))
 
```

We used this wrangled data to create the bar graph shown here. It displays the total amount of money contributed in favor of and in opposition of: challengers, incumbents, and open seat canidates in 2012. 
```{r}
library(ggthemes)

labels<-c(agcan="Against Candidate", forcan="For Candidate")

contrib_foragplot <- ggplot(cand_contrib, aes(x = cand_ici), transaction_amt) + geom_bar( fill = "darkgoldenrod1") + facet_wrap(~for_cand, labeller=labeller(for_cand=labels)) +
  theme_economist()+theme(legend.position = "none",
        axis.title.x = element_blank(), 
        axis.title.y = element_text(size=18, family="Andale Mono"),
         plot.title=element_text(size=20, family = "Andale Mono"), 
        axis.text.x = element_text(family = "Andale Mono"), 
        axis.text.y = element_text(family = "Andale Mono"))+
  scale_x_discrete(labels=c("Challenger", "Incumbent", "Open Seat"))+
  ylab("Total contributions (U.S. $)")+
  labs(title="  Campaign Contributions in 2012")

contrib_foragplot
```

This displays that the most money was spent against challengers, and the least money was spent against canidates where there was no incumbent. However, the variation between the three categories is small. This might be due to...

We can also see that in 2012, the most money was spent for incumbents by far. Incumbents might have more support because they have had more time to gain the trust of their constituents. (this isn't saying what I'm trying to say lol)

Generalized function: We tried to do more than one year, but failed because...2012
```{r}
contrib_any_year<-function(year){
  cand_contrib_yearly<-cand_contrib%>%
  filter(cand_election_yr == year)%>% 
  na.omit() %>%
  group_by(transaction_type)%>%
  mutate(for_cand = ifelse(transaction_type == "24K"|transaction_type ==  "24E", "forcan","agcan"))
  
 cand_contrib_yearly
  
  contrib_foragplot <- ggplot(cand_contrib_yearly, aes(x = cand_ici), transaction_amt) + geom_bar( fill = "darkgoldenrod1") + facet_wrap(~for_cand, labeller=labeller(for_cand=labels)) +
  theme_economist()+theme(legend.position = "none",
        axis.title.x = element_blank(), 
        axis.title.y = element_text(size=18, family="Andale Mono"),
         plot.title=element_text(size=20, family = "Andale Mono"), 
        axis.text.x = element_text(family = "Andale Mono"), 
        axis.text.y = element_text(family = "Andale Mono"))+
  scale_x_discrete(labels=c("Challenger", "Incumbent", "Open Seat"))+
  ylab("Total contributions (U.S. $)")

contrib_foragplot
}
```


```{r}
years<-c(1990, 1992, 1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 2012, 2014)

contrib_any_year(2012)
```


```{r, message=FALSE, warning=FALSE, eval=FALSE}
contrib_for <- cand_contrib %>%
  filter(!is.na("for_cand"))%>%
  filter(for_cand == "forcan")
 

contrib_ag <- cand_contrib %>%
  filter(for_cand == "agcan")%>%
  filter(!is.na("ag_cand"))
```

```{r, warning=FALSE, eval=FALSE}

boxplot <- ggplot(cand_contrib, aes(for_cand, log10(transaction_amt))) + geom_boxplot(color = "cornflowerblue")
boxplot
```

```{r, warning=FALSE, eval=FALSE}
contrib_forplot <- ggplot(contrib_for, aes(x = cand_ici, log10(transaction_amt))) + geom_boxplot()

contrib_forplot

contrib_foragplot <- ggplot(cand_contrib, aes(x = cand_ici, log10(transaction_amt))) + geom_boxplot(color = "cornflowerblue") + facet_wrap(~for_cand) 

contrib_foragplot
```

