---
title: "Mini-Project 2: Do Incumbents Receive More Money?"
author: "Grace Hartley, Amrita Acharya, Callie Slevin"
date: "2019-03-24"
output: 
  html_document:
      code_folding: hide
      theme: cosmo
---

^[https://github.com/hartlegr/sds192-mp2]

```{r, message=FALSE, warning = FALSE}
library(tidyverse)
load("candidates.rda")
load("contributions.rda")
devtools::install_github("benmarwick/wordcountaddin")

wordcountaddin::word_count("mp2.rmd")

```

We were interested in the sum of contributions made for and against the campaigns of different statuses of candidates (challenger, incumbent, and open seat candidates).

First, we joined the candidates table and the contributions table using the "cand_id."
```{r}
cand_contrib <- candidates%>% 
  left_join(contributions, "cand_id")
```


```{r, eval=FALSE, echo=FALSE}
library(Stat2Data)
tally(~transaction_type, data = cand_contrib)
```

Then, we filtered for the year 2012, while removing the rows with NA data. This provided us with all the contributions for candidates who last ran for office in 2012. We grouped the contributions by transaction type, and then mutated to add a column called 'for_cand'. This column provided information on whether the contribution was in favor of or in opposition of the candidate listed, listing either 'forcan' or 'agcan' respectively. 
```{r}
cand_contrib <- cand_contrib%>%
  filter(cand_election_yr == 2012)%>% 
  na.omit() %>%
  group_by(transaction_type)%>%
  mutate(for_cand = ifelse(transaction_type == "24K"|transaction_type ==  "24E", "forcan","agcan"))
 
```

We used this wrangled data to create the bar graph shown here. It displays the total amount of money contributed in favor of and in opposition of: challengers, incumbents, and open seat candidates in 2012. 
```{r}
library(ggthemes)

labels<-c(agcan="Against Candidate", forcan="For Candidate")

contrib_foragplot <- ggplot(cand_contrib, aes(x = cand_ici), transaction_amt) + geom_bar( fill = "darkgoldenrod1") + facet_wrap(~for_cand, labeller=labeller(for_cand=labels)) +
  theme_economist()+theme(legend.position = "none",
        axis.title.x = element_blank(), 
        axis.title.y = element_text(size=18, family="Andale Mono"),
         plot.title=element_text(size=20, family = "Andale Mono"), 
        axis.text.x = element_text(family = "Andale Mono"), 
        axis.text.y = element_text(family = "Andale Mono"))+
  scale_x_discrete(labels=c("Challenger", "Incumbent", "Open Seat"))+
  ylab("Total contributions (U.S. $)")+
  labs(title="  Campaign Contributions in 2012")

contrib_foragplot
```

>The plot shows that there is not a large difference in the money spent for and against the campaigns of both challengers and open seat candidates; the money spent for each group was about the same as the money spent against each group. 

Incumbents, however, received almost 10 times more money than was spent against them.
This is likely because incumbents have more support from their constituents (or organizations like West Virginians for Life or Holiday Inn Express) because they  have existing connections. Though any politician will have both supporters and opposers, an incumbent has had an oppurtunity to show the public how well they can do their job. A challenger relies more on their word and other (possibly less applicable) experiences. More support and exposure yields more money, evidently.

##For or Against?
Though the amounts were quite different between some of the groups, the distribution of the money between for and against was the same for each type of candidate. The total amount of money contributed for candidates was more than the amount contributed against, indicating that those donating money were more concerned with supporting their candidate of choice instead of undermining their opposition. 

Regarding money spent against any type of candidate, the least was spent against open seat campaigns, and the most was spent against challengers. This could be because candidates who have the most support (and therefore money to spare) are probably the incumbents. Some of their supporters who want to see them stay in office will focus on spending money against the challengers. In open seat elections, both candidates are more politically inexperienced (presumably), and the support from the public is not skewed one way or the other. Considering the sitution, contributions might seen as more valuable in favor of candidates, while oppositional contributions are more of a luxury.

##Function
Here is the generalized function that we created. It executes the same data wrangling and produces the same bar graph as above, but with any year. Our hope was to be able to map the function onto every year since 1990 to observe the trends in contributions for and against the three types of candidates.
```{r}
contrib_any_year<-function(year){
  cand_contrib_yearly<-cand_contrib%>%
  filter(cand_election_yr == year)%>% 
  na.omit() %>%
  group_by(transaction_type)%>%
  mutate(for_cand = ifelse(transaction_type == "24K"|transaction_type ==  "24E", "forcan","agcan"))
  
 cand_contrib_yearly
  
  contrib_foragplot <- ggplot(cand_contrib_yearly, aes(x = cand_ici), transaction_amt) + geom_bar( fill = "darkgoldenrod1") + facet_wrap(~for_cand, labeller=labeller(for_cand=labels)) +
  theme_economist()+theme(legend.position = "none",
        axis.title.x = element_blank(), 
        axis.title.y = element_text(size=18, family="Andale Mono"),
         plot.title=element_text(size=20, family = "Andale Mono"), 
        axis.text.x = element_text(family = "Andale Mono"), 
        axis.text.y = element_text(family = "Andale Mono"))+
  scale_x_discrete(labels=c("Challenger", "Incumbent", "Open Seat"))+
  ylab("Total contributions (U.S. $)")

contrib_foragplot
}
```

Though the function works, we weren't able to apply it to any year besides 2012 (shown below). The dataset provided on information on candidates in the *last*  year that they ran for office, so an incumbent who has been in office since the 1990s would only show up in 2012. This results in a lot of the data needed for this function being not available. 

```{r}
years<-c(1990, 1992, 1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 2012, 2014)

contrib_any_year(2012)
```


```{r, message=FALSE, warning=FALSE, eval=FALSE, echo=FALSE}
contrib_for <- cand_contrib %>%
  filter(!is.na("for_cand"))%>%
  filter(for_cand == "forcan")
 

contrib_ag <- cand_contrib %>%
  filter(for_cand == "agcan")%>%
  filter(!is.na("ag_cand"))
```

```{r, warning=FALSE, eval=FALSE, echo=FALSE}

boxplot <- ggplot(cand_contrib, aes(for_cand, log10(transaction_amt))) + geom_boxplot(color = "cornflowerblue")
boxplot
```

```{r, warning=FALSE, eval=FALSE, echo=FALSE}
contrib_forplot <- ggplot(contrib_for, aes(x = cand_ici, log10(transaction_amt))) + geom_boxplot()

contrib_forplot

contrib_foragplot <- ggplot(cand_contrib, aes(x = cand_ici, log10(transaction_amt))) + geom_boxplot(color = "cornflowerblue") + facet_wrap(~for_cand) 

contrib_foragplot
```

